\textbf{Title}

### 2015-10-26 paper/draft_3_no_fig.tex
added draft with no figures

\begin{figure}[ht]
\centering
	   \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_gaus_homo_comp.pdf}
        \caption{}
        \label{fig:gaus_homo_comp}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_gaus_homo_smash.pdf}
        \caption{}
        \label{fig:gaus_homo_smash}
    \end{subfigure}
    \caption{Violin plots of MISEs for various methods. Dashed green line indicates the median MISE for SMASH. a) compares SMASH against two popular and accurate denoising methods: TI-thresholding and Ebayesthresh, and b) demonstrates the performance gain of SMASH when it is provided the true variance.}
    \label{fig:gaus_homo}
\end{figure}

Having shown that SMASH performs well in the homoskedastic case without any explicit assumptions on the Gaussian noise, we now demonstrate its performance gain when the errors are heteroskedastic. Figure \ref{fig:simple_eg} demonstrates that SMASH is able to better capture the true signal than TI-thresh (with variance estimated by running median absolute deviation (RMAD)) on a simple simulation. For the sake of presentation, figure \ref{fig:gaus_hetero} highlights only the results from SMASH, TI-thresh (with various choices of variance estimation) and EbayesThresh for the ``Spikes'' mean function, with the ``Clipped Blocks'' variance function, a SNR of 3 and a sample size of 1024. Nevertheless, similar results hold in general for different mean and variance functions, SNRs, and sample sizes. 
\begin{figure}[ht]
\centering
	\includegraphics[width=\textwidth]{simple_eg.pdf}
	\caption{Simple illustration of heteroskedastic Gaussian denoising. Top left plot shows the mean function, top right plot shows the variance function, bottom left plot shows a sample dataset, and bottom right plot shows the estimated mean functions from SMASH and TI-thresh against the true mean function.}
	\label{fig:simple_eg}
\end{figure}

\begin{figure}[ht]
\centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_gaus_hetero_ti.pdf}
        \caption{}
        \label{fig:gaus_hetero_ti}
    \end{subfigure}
		\hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_gaus_hetero_homo.pdf}
        \caption{}
        \label{fig:gaus_hetero_homo}
    \end{subfigure}
		\hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_gaus_hetero_smash.pdf}
        \caption{}
        \label{fig:gaus_hetero_smash}
    \end{subfigure}
    \caption{Violin plots of MISEs for various methods. Dashed green line indicates the median MISE for SMASH. a) compares SMASH against TI-thresh with variance estimated by RMAD, variance estimated by SMASH, and the true variance. b) demonstrates the importance of taking heteroskedasticity into account, and c) reveals the performance gain of SMASH when it is provided the true variance.}
    \label{fig:gaus_hetero}
\end{figure}

@@

Since our method also provides variance estimates, one could potentially perform similar assessments for different variance functions as with mean functions. In this particular study, we compare our method against the only joint mean and variance estimation procedure for which software is easily available. Specifically, we consider the Mean Field Variational Bayes (MFVB) methodology developed by Menictas \& Wand (2014) for heteroskedastic Gaussian regression. Since MFVB is splines-based (and hence less adaptive locally), we chose the smooth mean and standard deviation functions (A) from Menictas \& Wand (2014) as our test functions, plotted in Figure \ref{fig:mfvb_fn} and denoted by $m(\cdot)$ and $sd(\cdot)$ respectively. We then considered two different simulation scenarios:
\begin{enumerate}
\item $n=500$ independent $(X_i,Y_i)$ pairs were generated. The $X_i$'s were distributed as Uniform(0,1), and the $Y_i$'s were distributed as $N(m(x_i),sd(x_i))$. The performance of a given method is measured by the standard MSE evaluated at 201 equally spaced points on $(X_{min},X_{max})$ for both the mean and the standard deviation functions.
\item $n=1024$ independent $(X_i,Y_i)$ pairs were generated. The $X_i$'s were deterministic and equally spaced on (0,1), while the $Y_i$'s were distributed as $N(m(x_i),sd(x_i))$. The performance of a given method is measured by the MSE evaluated at the 1024 $X_i$'s for both the mean and the standard deviation functions.
\end{enumerate}
\begin{figure}[ht]
\centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{mfvb_mean.pdf}
        \caption{}
        \label{fig:mfvb_mean}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{mfvb_var.pdf}
        \caption{}
        \label{fig:mfvb_var}
    \end{subfigure}
    \caption{Mean (a) and variance (b) functions used in the simulation study comparing SMASH against MFVB.}
    \label{fig:mfvb_fn}
\end{figure}
In the first scenario, the number of data points is not a power of two, nor are the point equally spaced. To deal with these complications, we adapted and modified the standard symmetric extension procedure commonly used in wavelet settings. We first mirrored the data about the right edge and extract the first $2^{\lfloor\log_2(2n)\rfloor}$ sample points. This ensures that the number of data points in the new ``dataset'' is a power of two, and the mean curve would be continuous at the right edge. To further ensure that the input to the Gaussian denoising method is periodic, we then reflected the new dataset about the right edge and used this as the final input. To obtain the original mean and variance functions, we extracted the first $n$ points from the outputs (mean and variance) of our denoising technique. Since the data points are not evenly spaced, we took the simplest approach and applied our method treating the observations as if they were evenly spaced. This approach is not only intuitively appealing, but can also be considered a formal treatment of unequally spaced data in traditional wavelet settings (see Sardy et al. (1999)). Evaluation of MSE at the 201 equally spaced points is then based on simple linear interpolation between the estimated points. Tables \ref{table:mfvb_comp} display the MSEs over 100 independent runs for each scenario.
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
\hline
& \multicolumn{2}{c}{Scenario 1}&\multicolumn{2}{c}{Scenario 2}\\
\cline{2-5}
& mean & sd & mean & sd \\
\hline
MFVB & 0.0330 & 0.0199 & 0.0172 & 0.0085 \\
SMASH & 0.0334 & 0.0187 & 0.0158 & 0.0065 \\
\hline
\end{tabular}
\caption{MSEs of MFVB and SMASH for two simulation scenarios}
\label{table:mfvb_comp}
\end{table}

@@

Thus far, we have demonstrated that our method does a good job of mean and variance estimation for a variety of situations in the Gaussian case. As such, we now turn our attention to the Poisson case. For this simulation study, we considered different test functions and Poisson intensities for a given sample size of $n=1024$ as in Timmermann \& Nowak (1999) and Fryzlewicz \& Nason (2004), over 100 independent runs. Figure \ref{fig:pois_sim} compares the MISE of SMASH with Haar-Fisz (Fryzlewicz \& Nason (2004)) and BMSM (Kolaczyk (1999b)), which are two popular and accurate Poisson denoising techniques, for the ``Bursts'' function with (min,max) intensities of (0.01,3) and (1/8,8), which are of primary interest. Complete results, including a (min,max) intensity of (1/128, 128), are included in appendix \textcolor{red}{where?}. From figure \ref{fig:pois_sim}, it is clear that SMASH outperforms both methods, and the claim holds true in general. As an exception, Haar-Fisz outperforms both SMASH and BMSM for certain test functions with a (min,max) intensity of (1/128,128) when the asmyptotic variance of 1 is assumed in the Gaussian wavelet thresholding stage of the Haar-Fisz algorithm, but underperforms when the variance is estimated from the data instead. Besides the assumption of unit variance, the inconsistent performance of Haar-Fisz could also be attributed to the choice of the primary resolution level used, which can substantially affect the performance of Haar-Fisz. Here we analyzed primary resolution levels of 4, 5, 6 and 7. On the other hand, SMASH outperforms BMSM consistently, even though our method is based on the same likelihood factorization used in the latter. We can thus conclude that the choice of ASH as the shrinkage procedure (with the necessary likelihood approximations) is superior to that used in BMSM. As will be discussed later, using the formulation in SMASH also makes it easily extensible to multiple samples in the context of a (generalized) linear model.
\begin{figure}
\centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_pois_1.pdf}
        \caption{}
        \label{fig:pois_1}
    \end{subfigure}
		\hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_pois_8.pdf}
        \caption{}
        \label{fig:pois_8}
    \end{subfigure}
    \caption{Violin plots of MISEs for various methods. a) corresponds to a (min,max) intensity of (0.01,3), and b) corresponds to a (min,max) intensity of (1/8,8)}
    \label{fig:pois_sim}
\end{figure}

@@

\begin{figure}
\centering
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures_treasury_a.pdf}
        \caption{}
        \label{fig:tb_raw_data}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures_treasury_b.pdf}
        \caption{}
        \label{fig:tb_res}
    \end{subfigure}
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures_treasury_c.pdf}
        \caption{}
        \label{fig:tb_mean}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures_treasury_d.pdf}
        \caption{}
        \label{fig:tb_var}
    \end{subfigure}
    \caption{Analysis of yields from three-month Treasury Bills. (a) Raw data of interest rates as a function of time. (b) Residuals $Y_t$ from fitting an AR(5) model to the data against $X_t\equiv T_{t-1}$. The red curve is the estimated mean curve from SMASH. (c) Plot of a zoomed-in version of the estimated mean curve, with approximate 95\% credible bands. (d) The estimated conditional variance curve, with approximate 95\% credible bands.}
\end{figure}

@@

\begin{figure}
\centering
    \begin{subfigure}[b]{0.85\textwidth}
        \centering
        \includegraphics[width=\textwidth]{peaks_comp_a.pdf}
        \caption{}
        \label{fig:seq_peak_data}
    \end{subfigure}
    \begin{subfigure}[b]{0.85\textwidth}
        \centering
        \includegraphics[width=\textwidth]{peaks_comp_b.pdf}
        \caption{}
        \label{fig:seq_peak_est}
    \end{subfigure}
    \caption{Analysis of reads taken from the GM12878 cell line, spanning position 880001 to 1011072 on chromosome 1. (a) Raw sequencing counts. (b) Estimated intensity function from SMASH (black solid line) and location of peaks called by MACS (red markers beneath the estimated intensity).}
    \label{fig:seq_peak}
\end{figure}

>>>>

%\begin{figure}[ht]
%\centering
	   %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{violin_gaus_homo_comp.pdf}
        %\caption{}
        %\label{fig:gaus_homo_comp}
    %\end{subfigure}
    %\hfill
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{violin_gaus_homo_smash.pdf}
        %\caption{}
        %\label{fig:gaus_homo_smash}
    %\end{subfigure}
    %\caption{Violin plots of MISEs for various methods. Dashed green line indicates the median MISE for SMASH. a) compares SMASH against two popular and accurate denoising methods: TI-thresholding and Ebayesthresh, and b) demonstrates the performance gain of SMASH when it is provided the true variance.}
    %\label{fig:gaus_homo}
%\end{figure}

Having shown that SMASH performs well in the homoskedastic case without any explicit assumptions on the Gaussian noise, we now demonstrate its performance gain when the errors are heteroskedastic. Figure \ref{fig:simple_eg} demonstrates that SMASH is able to better capture the true signal than TI-thresh (with variance estimated by running median absolute deviation (RMAD)) on a simple simulation. For the sake of presentation, figure \ref{fig:gaus_hetero} highlights only the results from SMASH, TI-thresh (with various choices of variance estimation) and EbayesThresh for the ``Spikes'' mean function, with the ``Clipped Blocks'' variance function, a SNR of 3 and a sample size of 1024. Nevertheless, similar results hold in general for different mean and variance functions, SNRs, and sample sizes. 
%\begin{figure}[ht]
%\centering
	%\includegraphics[width=\textwidth]{simple_eg.pdf}
	%\caption{Simple illustration of heteroskedastic Gaussian denoising. Top left plot shows the mean function, top right plot shows the variance function, bottom left plot shows a sample dataset, and bottom right plot shows the estimated mean functions from SMASH and TI-thresh against the true mean function.}
	%\label{fig:simple_eg}
%\end{figure}    
%
%\begin{figure}[ht]
%\centering
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{violin_gaus_hetero_ti.pdf}
        %\caption{}
        %\label{fig:gaus_hetero_ti}
    %\end{subfigure}
		%\hfill
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{violin_gaus_hetero_homo.pdf}
        %\caption{}
        %\label{fig:gaus_hetero_homo}
    %\end{subfigure}
		%\hfill
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{violin_gaus_hetero_smash.pdf}
        %\caption{}
        %\label{fig:gaus_hetero_smash}
    %\end{subfigure}
    %\caption{Violin plots of MISEs for various methods. Dashed green line indicates the median MISE for SMASH. a) compares SMASH against TI-thresh with variance estimated by RMAD, variance estimated by SMASH, and the true variance. b) demonstrates the importance of taking heteroskedasticity into account, and c) reveals the performance gain of SMASH when it is provided the true variance.}
    %\label{fig:gaus_hetero}
%\end{figure}

@@

Since our method also provides variance estimates, one could potentially perform similar assessments for different variance functions as with mean functions. In this particular study, we compare our method against the only joint mean and variance estimation procedure for which software is easily available. Specifically, we consider the Mean Field Variational Bayes (MFVB) methodology developed by Menictas \& Wand (2014) for heteroskedastic Gaussian regression. Since MFVB is splines-based (and hence less adaptive locally), we chose the smooth mean and standard deviation functions (A) from Menictas \& Wand (2014) as our test functions, plotted in Figure \ref{fig:mfvb_fn} and denoted by $m(\cdot)$ and $sd(\cdot)$ respectively. We then considered two different simulation scenarios:
\begin{enumerate}
\item $n=500$ independent $(X_i,Y_i)$ pairs were generated. The $X_i$'s were distributed as Uniform(0,1), and the $Y_i$'s were distributed as $N(m(x_i),sd(x_i))$. The performance of a given method is measured by the standard MSE evaluated at 201 equally spaced points on $(X_{min},X_{max})$ for both the mean and the standard deviation functions.
\item $n=1024$ independent $(X_i,Y_i)$ pairs were generated. The $X_i$'s were deterministic and equally spaced on (0,1), while the $Y_i$'s were distributed as $N(m(x_i),sd(x_i))$. The performance of a given method is measured by the MSE evaluated at the 1024 $X_i$'s for both the mean and the standard deviation functions.
\end{enumerate}
%\begin{figure}[ht]
%\centering
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{mfvb_mean.pdf}
        %\caption{}
        %\label{fig:mfvb_mean}
    %\end{subfigure}
    %\hfill
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{mfvb_var.pdf}
        %\caption{}
        %\label{fig:mfvb_var}
    %\end{subfigure}
    %\caption{Mean (a) and variance (b) functions used in the simulation study comparing SMASH against MFVB.}
    %\label{fig:mfvb_fn}
%\end{figure}
In the first scenario, the number of data points is not a power of two, nor are the point equally spaced. To deal with these complications, we adapted and modified the standard symmetric extension procedure commonly used in wavelet settings. We first mirrored the data about the right edge and extract the first $2^{\lfloor\log_2(2n)\rfloor}$ sample points. This ensures that the number of data points in the new ``dataset'' is a power of two, and the mean curve would be continuous at the right edge. To further ensure that the input to the Gaussian denoising method is periodic, we then reflected the new dataset about the right edge and used this as the final input. To obtain the original mean and variance functions, we extracted the first $n$ points from the outputs (mean and variance) of our denoising technique. Since the data points are not evenly spaced, we took the simplest approach and applied our method treating the observations as if they were evenly spaced. This approach is not only intuitively appealing, but can also be considered a formal treatment of unequally spaced data in traditional wavelet settings (see Sardy et al. (1999)). Evaluation of MSE at the 201 equally spaced points is then based on simple linear interpolation between the estimated points. Tables \ref{table:mfvb_comp} display the MSEs over 100 independent runs for each scenario.
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
\hline
& \multicolumn{2}{c}{Scenario 1}&\multicolumn{2}{c}{Scenario 2}\\
\cline{2-5}
& mean & sd & mean & sd \\
\hline
MFVB & 0.0330 & 0.0199 & 0.0172 & 0.0085 \\
SMASH & 0.0334 & 0.0187 & 0.0158 & 0.0065 \\
\hline
\end{tabular}
\caption{MSEs of MFVB and SMASH for two simulation scenarios}
\label{table:mfvb_comp}
\end{table}

@@

Thus far, we have demonstrated that our method does a good job of mean and variance estimation for a variety of situations in the Gaussian case. As such, we now turn our attention to the Poisson case. For this simulation study, we considered different test functions and Poisson intensities for a given sample size of $n=1024$ as in Timmermann \& Nowak (1999) and Fryzlewicz \& Nason (2004), over 100 independent runs. Figure \ref{fig:pois_sim} compares the MISE of SMASH with Haar-Fisz (Fryzlewicz \& Nason (2004)) and BMSM (Kolaczyk (1999b)), which are two popular and accurate Poisson denoising techniques, for the ``Bursts'' function with (min,max) intensities of (0.01,3) and (1/8,8), which are of primary interest. Complete results, including a (min,max) intensity of (1/128, 128), are included in appendix \textcolor{red}{where?}. From figure \ref{fig:pois_sim}, it is clear that SMASH outperforms both methods, and the claim holds true in general. As an exception, Haar-Fisz outperforms both SMASH and BMSM for certain test functions with a (min,max) intensity of (1/128,128) when the asmyptotic variance of 1 is assumed in the Gaussian wavelet thresholding stage of the Haar-Fisz algorithm, but underperforms when the variance is estimated from the data instead. Besides the assumption of unit variance, the inconsistent performance of Haar-Fisz could also be attributed to the choice of the primary resolution level used, which can substantially affect the performance of Haar-Fisz. Here we analyzed primary resolution levels of 4, 5, 6 and 7. On the other hand, SMASH outperforms BMSM consistently, even though our method is based on the same likelihood factorization used in the latter. We can thus conclude that the choice of ASH as the shrinkage procedure (with the necessary likelihood approximations) is superior to that used in BMSM. As will be discussed later, using the formulation in SMASH also makes it easily extensible to multiple samples in the context of a (generalized) linear model.
%\begin{figure}
%\centering
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{violin_pois_1.pdf}
        %\caption{}
        %\label{fig:pois_1}
    %\end{subfigure}
		%\hfill
    %\begin{subfigure}[b]{0.45\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{violin_pois_8.pdf}
        %\caption{}
        %\label{fig:pois_8}
    %\end{subfigure}
    %\caption{Violin plots of MISEs for various methods. a) corresponds to a (min,max) intensity of (0.01,3), and b) corresponds to a (min,max) intensity of (1/8,8)}
    %\label{fig:pois_sim}
%\end{figure}

@@

%\begin{figure}
%\centering
    %\begin{subfigure}[b]{0.4\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{figures_treasury_a.pdf}
        %\caption{}
        %\label{fig:tb_raw_data}
    %\end{subfigure}
    %\hfill
    %\begin{subfigure}[b]{0.4\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{figures_treasury_b.pdf}
        %\caption{}
        %\label{fig:tb_res}
    %\end{subfigure}
    %\begin{subfigure}[b]{0.4\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{figures_treasury_c.pdf}
        %\caption{}
        %\label{fig:tb_mean}
    %\end{subfigure}
    %\hfill
    %\begin{subfigure}[b]{0.4\textwidth}
        %\centering
        %\includegraphics[width=\textwidth]{figures_treasury_d.pdf}
        %\caption{}
        %\label{fig:tb_var}
    %\end{subfigure}
    %\caption{Analysis of yields from three-month Treasury Bills. (a) Raw data of interest rates as a function of time. (b) Residuals $Y_t$ from fitting an AR(5) model to the data against $X_t\equiv T_{t-1}$. The red curve is the estimated mean curve from SMASH. (c) Plot of a zoomed-in version of the estimated mean curve, with approximate 95\% credible bands. (d) The estimated conditional variance curve, with approximate 95\% credible bands.}
%\end{figure}

@@

%\begin{figure}
    %\begin{subfigure}[b]{0.85\textwidth}
        %\includegraphics[width=\textwidth]{peaks_comp_a.pdf}
        %\caption{}
        %\label{fig:seq_peak_data}
    %\end{subfigure}
    %\begin{subfigure}[b]{0.85\textwidth}
        %\includegraphics[width=\textwidth]{peaks_comp_b.pdf}
        %\caption{}
        %\label{fig:seq_peak_est}
    %\end{subfigure}
    %\caption{Analysis of reads taken from the GM12878 cell line, spanning position 880001 to 1011072 on chromosome 1. (a) Raw sequencing counts. (b) Estimated intensity function from SMASH (black solid line) and location of peaks called by MACS (red markers beneath the estimated intensity).}
    %\label{fig:seq_peak}
